Regional Inequalities and economic growth; cross section and times series analysis

### 2023-11-20 MSB104_GR_1_Final_Assignment_research_article.tex
:D

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\KOMAoption{captions}{tableheading}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Regional Inequalities and economic growth; cross section and times series analysis},
  pdfauthor={Anine Therese Karlsen \& Mona Lisa Jones},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{document}
\maketitle
\begin{abstract}
This assignment aims to acquire, process, and analyze sub-national GDP
and population data for a subset of European countries. Calculate GDP
per capita and explore regional inequity using various descriptive
statistics and visualizations.Furthermore examining the relationship
between regional development and inequality, employing cross sectional
estimation techniques for the year 2010.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, sharp corners, frame hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, interior hidden, enhanced]}{\end{tcolorbox}}\fi

@@

While national GDP and GDP per capita are vital indicators of a
country's aggregate economic health, they do not shed light on how
wealth or income is distributed among its residents. A high national GDP
can, paradoxically, coexist with pockets of regional deprivation
(\textbf{lessmann2017?}).

The truth of this statement becomes more evident when taking a closer
look at sub-national data. Regional wealth disparities are of prime
concern, especially when crafting policies for equitable growth
(\textbf{lessmann2017?}). A country's macro-level prosperity does not
automatically guarantee that all its regions partake equally in this
wealth. By studying smaller regions within a country, it is possible to
get a more nuanced narrative about the state of regional economic
disparities (\textbf{lessmann2017?}).

Using time series data and cross sectional observations we investigate
the GDP per capita and GINI trends in Portugal, France, Hungary,
Slovenia and Denmark in the time period 2000 to 2020. The aims is to
acquire, process, and analyze sub-national GDP and population data for a
subset of European countries. Calculate GDP per capita and explore
regional inequity using various descriptive statistics and
visualizations

\hypertarget{literature-review}{%
\section{Literature review}\label{literature-review}}

Regional growth and inequality have been important topics of research in
Europe for many years. In recent decades, the European Union has
implemented policies aimed at reducing regional disparities and
promoting economic growth across its member states. Furthermore, despite
these efforts, significant disparities in income and economic
development persist across regions in Europe (\textbf{lessmann2017?}).

When it comes to economic growth, several studies refer to numerous
influencing factors. (\textbf{gennaioli2014?}) and
(\textbf{crescenzi2012?}) states that regional growth, much like
national growth, is significantly influenced by geography and human
capital. Geography impacts trade, resource availability, and
susceptibility to natural events. Human capital, encompassing skills and
knowledge, drives innovation and productivity. Furthermore,
(\textbf{gennaioli2014?}) mentions that wealthier nations like Denmark
in our case, tend to experience faster regional convergence, suggesting
national prosperity aids regional growth (\textbf{gennaioli2014?}).
Convergence implies that poorer regions grow faster than richer ones,
gradually reducing disparities in wealth and income. Also referred to as
beta convergence in (\textbf{lessmann2017?}) paper.

In the context of Europe, and our selected countries we might expect to
see a wide variety of regional economic performances, especially within
Hungary and Slovakia and between the western European countries. Capital
market regulations also play a fundamental role, with effective
frameworks promoting quicker regional convergence
(\textbf{gennaioli2014?}). Taken this in to consideration we might
expect a faster growth rate after the joining of the EU in 2003 and
2004. That said the opening of the free labor market might also
influence the migration of labor in Slovakia and Hungary. Not to mention
the economic crisis hitting shortly after our eastern European countries
joined the union.

Whilst their recent EU membership might have posed certain challenges
for Eastern European countries during the financial crisis, it also
offered tools and resources that aided in their recovery in Slovakia and
Hungary. As to the western Europeans such as Denmark, Portugal and
France we expect to see a sharp decline in the economic growth.
Likewise, with the recent covid19 pandemic in terms of financial
aspects.

Regarding the effects of the economic crisis, another paper notes that
it led to a significant increase in regional unemployment rates across
Europe. Reporting some regions experiencing unemployment rates as high
as 30\% (\textbf{iammarino2019?}). The authors also discuss how the
crisis exposed the vulnerability of regions that were heavily reliant on
a single industry or sector, and how sheltered economies were better
able to weather the storm (\textbf{iammarino2019?}). Another affect that
surprised us was that the economic crisis led to increase inequality.
The study ``\emph{the effects of financial crisis on regional
inequality}'' indicate any type of financial crisis results in higher
income inequality (\textbf{nguyen2022?}).

(\textbf{lessmann2017?}) regional inequality study address the
importance of studying regional inequality. In particular, its potential
consequences, such as political tensions that can undermine social and
political stability (\textbf{lessmann2017?}). The paper also discusses
the relationship between regional inequality and personal income
inequality and conflict risk. Furthermore, the paper provides us with
insights into the inverted U-shaped relationship between regional
inequality and the level of economic development in different country
groups (\textbf{lessmann2017?}). The inverted U-shape also referred to
as Kuznets Curve suggests that as an economy develops, regional
inequality initially rises, reaches a peak, and then starts to decline.
(\textbf{lessmann2017?}) paper also finds that there is an N-shaped
relationship between regional inequality and economic development, which
means that regional inequality increases again after the inverted
U-shaped pattern has been completed (\textbf{lessmann2017?}).

Lastly, (\textbf{iammarino2019?}) discusses regional inequalities in
European countries. It presents evidence that inter-regional inequality
has been increasing in the European Union since the turn of the
millennium. The authors argue that this is due to the existence of
several groups of regional economies in Europe that are structurally
very different from one another (\textbf{iammarino2019?}). Additionally,
the authors argue that there are some countries in the EU that are more
evenly developed than others, and that a map of under performance or
over-performance means less in a high-income but evenly developed
country, such as in Denmark in our case (\textbf{iammarino2019?}). On
the other hand, France has been a subject of discussion and policymaking
for many years, regarding its pockets of extreme wealth. (finner ikke
kilde)

\hypertarget{data}{%
\section{Data}\label{data}}

Trough Eurostat, we download the datasets nama\_10r\_3gdp and
demo\_r\_pjanggr3 as csv files and filted the data by our preferences
before downloading it. Furthermore, we filtered the dataset by choosing
the years 2000 to 2020. Then, we selected the NUTS 3 region for
Portugal, France, Hungary, Slovakia and Denmark. Finally, we specified
the data to be in million Euro.

\hypertarget{gdp}{%
\subsubsection{GDP}\label{gdp}}

The GDP dataset provides insights into GDP at regional level using the
NUTS classification system. It furnishes GDP values in both current
prices and adjusted for inflation, with figures given in purchasing
power standards (PPS) to account for price level differences between
countries. The data is mostly structured by year and region
(\textbf{eurostat2023?}).

The GDP at market prices represents the final result of production
activities of resident producer units within a region or nation. It is
calculated as the sum of the gross value added across various
institutional sectors or industries. Furthermore, augmented by taxes and
reduced by subsidies on products (which are not allocated to specific
sectors or industries) (\textbf{eurostat2023c?}). This also balances out
in the total economy production account. In terms of methodology, while
national accounts compile GDP from the expenditure side, regional
accounts don't adopt this approach due to the complexities of accurately
mapping inter-regional flows of goods and services.

The different measures for the regional GDP are absolute figures in €
and Purchasing Power Standards (PPS), figures per inhabitant and
relative data compared to the EU Member States average
(\textbf{eurostat2023c?}).

\hypertarget{population}{%
\subsubsection{Population}\label{population}}

Eurostat's records annual population data with NUTS classification. Our
dataset includes information on births, deaths, net migration, and may
also include demographic information on age and gender. Displayed in a
year-by-region format, with yearly interval updates
(\textbf{eurostat2023a?}).

Eurostat's primary source for yearly demographic data at the regional
level stems from the Unified Demography (Unidemo) project. The project
covers 37 countries and is the central repository for demographic and
migration-related data (\textbf{eurostat2023a?}). As well as, specific
metrics gathered under UNIDEMO encompass population counts at the close
of the calendar year and events such as births and deaths occurring
within that year (\textbf{eurostat2023a?}). Additionally, data on
marriages, divorces, and migration flows are recorded .

For the purpose of this research, the demographic data references the
NUTS 2016 classification, which provides a detailed breakdown of the
European Union's territory (\textbf{eurostat2021?}).

\hypertarget{nuts-classification}{%
\subsubsection{NUTS classification}\label{nuts-classification}}

The Nomenclature of Territorial Units for Statistics (NUTS) offers a
stratified system to segment the economic territory of the EU and UK to
facilitate the consistent collection and harmonization of regional
statistics across Europe. The NUTS regions range from NUTS 0 Country
level to NUTS 3 small units such as municipalities level
(\textbf{eurostat2023a?}). ~

\hypertarget{econometric-approach}{%
\section{Econometric Approach}\label{econometric-approach}}

In this part we will report and interpret different types of essential
descriptive statistics. Measuring regional income inequality is
challenging due to heterogeneity of regions (\textbf{lessmann2017?}).
The number of regions in our data set varies largely in size and
population. Since the focus of this paper is purely growth and changes
in inequities over time, the variations of size and population density
becomes a minor issue because the country-level territorial
heterogeneity is fixed (\textbf{lessmann2017?}).

``Interest in income inequality has led to the development of several
ways of measuring it. Two types of measures are of interest in this
paper----static and dynamic. \emph{Static measures} provide a snapshot
(cross sectional) of these inequalities at a point of time whereas the
dynamic \emph{measures capture historical trends} (Paneldata)\emph{.}''
(\textbf{wooldridge2020?})

>>>>

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{soul}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage{caption}
\usepackage{graphicx}
\usepackage{siunitx}
\usepackage{hhline}
\usepackage{calc}
\usepackage{tabularx}
\usepackage{adjustbox}
\usepackage{hyperref}
\KOMAoption{captions}{tableheading}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Regional Inequalities and economic growth; cross section and times series analysis},
  pdfauthor={Anine Therese Karlsen \& Mona Lisa Jones},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{document}
\maketitle
\begin{abstract}
This assignment aims to acquire, process, and analyze sub-national GDP
and population data for a subset of European countries. Firstly
(assignment 1), Calculateing GDP per capita and exploreing regional
inequity using various descriptive statistics and
visualizations.Secondly (assignment 2), examining the relationship
between regional development and inequality, employing cross sectional
estimation techniques for the year 2010.Lastly exploring alternative
functional forms and implementing panel estimation to analzise the
relationship between regional development inequality
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, boxrule=0pt, breakable, enhanced, sharp corners, borderline west={3pt}{0pt}{shadecolor}, interior hidden]}{\end{tcolorbox}}\fi

@@

While national GDP and GDP per capita are vital indicators of a
country's aggregate economic health, they do not shed light on how
wealth or income is distributed among its residents. A high national GDP
can, paradoxically, coexist with pockets of regional deprivation
(Lessmann \& Seidel, 2017).

The truth of this statement becomes more evident when taking a closer
look at sub-national data. Regional wealth disparities are of prime
concern, especially when crafting policies for equitable growth
(Lessmann \& Seidel, 2017). A country's macro-level prosperity does not
automatically guarantee that all its regions partake equally in this
wealth. By studying smaller regions within a country, it is possible to
get a more nuanced narrative about the state of regional economic
disparities (Lessmann \& Seidel, 2017).

Using time series data and cross-sectional observations we investigate
Economic growth and Inequality trends in Portugal, France, Hungary,
Slovenia and Denmark in the time period 2000 to 2020. The aims are to
acquire, process, and analyse sub-national GDP, GINI and population data
for a subset of European countries.

The primary objective is to gain knowledge of econometric methods used
for science in business through our given topic. Subsequently,
econometric terms and definitions are explained in greater detail than
it normally would in similar papers. Using GDP per capita and the Gini
coefficient. Moreover, both are vital indicators in the field of
economics and are frequently used, in a wide array of studies.
Furthermore, to analyse and interpret the economic conditions and social
disparities of different regions or countries. Their analysis provides
crucial insights for policymakers, economists, and researchers in
shaping and evaluating the effectiveness of socio-economic policies.

The research question \textbf{\emph{``What are the regional economic
growth and inequality measurements from 2000 to 2020, in selected
European countries and how does other}} \textbf{\emph{determinants
affect the results?{}``}} is key in guiding our study. Setting a solid
foundation for a comprehensive study that not only measure our given
parameters, but also a critically examination of the role of
determinants such as transportation infrastructure and education.
Lessmann \& Seidel (2017) states that transportation costs play an
important role in agglomeration and income, and reduced transportation
costs as a determinant of regional inequality.

The sructure of our study is methodically designed to facilitate this
exploration. In part 1A and B we calculate GDP per capita and explore
regional inequality using various descriptive statistics and
visualizations such as panel data (\textbf{\emph{Assignment\_1}}).

In part 2A and B we use cross sectional method to examine the
relationship between regional development and inequality, employing
cross sectional estimation techniques for the year 2010
(\textbf{\emph{Assignment\_2}}).

Furthermore, exploring alternative functional forms and implementing
panel estimation to analzise the relationship between regional
development inequality (\textbf{\emph{Assignment\_3}}).

Lastly, it is worth noting potential limitations may arise due to the
use of ChatGPT and ChatGPT PDF, which have been utilized as tools to
gain understanding and extract knowledge from academic papers, as well
as to explain econometric terms and definitions. The text generated by
these tools has been cross-checked to the best of our knowledge with our
textbook and Andre Seidel, lecturer and author of `Regional Inequality,
Convergence, and Its Determinants -- A View from Outer Space' (2017).
Despite these precautions, there may have been instances where incorrect
information was provided, and some errors could have been overlooked.
The text has been meticulously rewritten to fit our style of writing,
ensuring consistency and accuracy throughout our study.

\hypertarget{literature-review-regional-economic-growth-inequality-and-determines-2000-2020}{%
\section{Literature Review: Regional Economic Growth, Inequality and
determines
(2000-2020)}\label{literature-review-regional-economic-growth-inequality-and-determines-2000-2020}}

Why some places grow and prosper is a fundamental question in social
science research. This question motivated Adam Smith's work, \emph{An
Inquiry into the Nature and Causes of the Wealth of Nations,} (Smith
\href{https://link-springer-com.galanga.hvl.no/article/10.1007/s11187-014-9574-4\#ref-CR32}{1863})
and has been a major influence of the \emph{new economic geography}
(Feldman, 2014)\emph{. Furthermore,} regional growth and inequality have
been important topics of research in Europe for many years. In recent
decades, the European Union implemented policies aimed at reducing
regional disparities and promoting economic growth across its member
states. Furthermore, despite these efforts, significant disparities in
income and economic development persist across regions in Europe
(Lessmann \& Seidel, 2017).

\textbf{Economic Growth and Regional Disparities}

Regional growth in Europe exhibits a complex pattern, influenced
significantly by factors such as geography, human capital, and economic
policies. Lessmann and Seidel's research offers insights into these
dynamics, highlighting the persistence of income disparities, across
regions despite overarching economic advancement (Lessmann \& Seidel,
2017). Wealthier nations like Denmark tend to experience quicker
regional convergence, a phenomenon supported by Gennaioli et al.'s
findings (Gennaioli et al., 2014).

\textbf{Financial crisis, and EU membership for Slovakia and Hungary}

The 2008 financial crisis significantly disrupted regional growth and
exacerbated income inequalities across Europe. Nguyen et al.~highlight
that financial turmoil typically leads to increased income disparity; a
trend observed globally (Nguyen, 2022).

Focusing on Europe, we expect varied regional economic performances,
especially when comparing Eastern European countries such as Hungary and
Slovakia to their Western counterparts. Given the role of capital market
regulations in promoting regional convergence, outlined by Gennaioli et
al., we anticipate a potential growth acceleration. In particular for
Eastern European countries following their EU accession in 2003 and 2004
(Gennaioli et al., 2014). However, the financial crisis and subsequent
labour market shifts may also have influenced these trends.

EU membership provided Slovakia and Hungary with resources for economic
recovery, while Western European countries, including Denmark, Portugal,
and France, likely experienced significant economic declines due to the
financial crisis, a pattern possibly mirrored in the recent COVID-19
pandemic. In particular Portugal as the crisis hit them harder. These
complexities highlight the importance of a detailed analysis to
understand regional economic growth and inequality in challenging times.

\textbf{Regional inequality}

Another affect that surprised us was that the economic crisis led to
increase inequality. The study ``\emph{the effects of financial crisis
on regional inequality}'' indicate any type of financial crisis results
in higher income inequality (Nguyen, 2022).

Lessmann \& Seidel (2017) regional inequality study address the
importance of studying these aspects. Its potential consequences, such
as political tensions that can undermine social and political stability
(Lessmann \& Seidel, 2017). The paper also discusses the relationship
between regional inequality and personal income inequality and conflict
risk.

\textbf{Determinants of regional inequality}

Geographic characteristics stand out as pivotal, with the natural and
topographical elements of a region significantly swaying income
distributions (Lessmann \& Seidel, 2017).

Urbanization emerges as a crucial determinant, where bustling, populous
around urban centers usually enjoy elevated income levels, contrasting
starkly with the often-stark disparities witnessed in more rural and
secluded areas (Lessmann \& Seidel, 2017).

\emph{Transport infrastructure and Education}

Transport infrastructure plays a central role in shaping regional
economic landscapes. Improved connectivity fosters economic activities,
reduces travel time, and enhances accessibility to markets and
resources. Studies indicates that regions with robust transport
infrastructure tend to attract more investments, generate employment,
and ultimately contribute to regional prosperity.

(Lessmann \& Seidel, 2017) also discusses the determinant of Education
in the context of human capital. The study mentions that human capital
is the most important determinant of differences in regional development
within countries. The study also notes that human capital creates
positive externalizes, which can spill over regional boundaries.
Moreover, the quality of human capital within countries, as measured by
the secondary-school enrollment rate, can decrease regional inequality
by facilitating regional spillovers and convergence. Lastly, promoting
internal migration.

\textbf{Expected Inverted U- and N shaped curves}

Furthermore, Lessmann \& Seidel (2017) provides us with insights into
the inverted U-shaped relationship between regional inequality and the
level of economic development in different country groups (Lessmann \&
Seidel, 2017). The inverted U-shape also referred to as Kuznets Curve
suggests that as an economy develops, regional inequality initially
rises, reaches a peak, and then starts to decline.

Lessmann \& Seidel (2017) paper also finds that there is an N-shaped
relationship between regional inequality and economic development, which
means that regional inequality increases again after the inverted
U-shaped pattern has been completed (Lessmann \& Seidel, 2017).

Iammarino et al. (2019) discusses regional inequalities in European
countries. It presents evidence that inter-regional inequality has been
increasing in the European Union since the turn of the millennium. The
authors argue that this is due to the existence of several groups of
regional economies in Europe that are structurally very different from
one another (Iammarino et al., 2019). Additionally, the authors argue
that there are some countries in the EU that are more evenly developed
than others, and that a map of under performance or over-performance
means less in a high-income but evenly developed country, such as in
Denmark in our case (Iammarino et al., 2019). On the other hand, France
has been a subject of discussion and policy making for many years,
regarding its well-known pockets of extreme wealth in regions such as
Cap d'Antibes.

Overall, it is evident that regional economic growth and inequality in
Europe from 2000 to 2020 have been influenced by a constellation of
factors, with transport infrastructure and education emerging as
significant determinants. Earlier studies provide us with a fundamental
understanding of these dynamics, while also highlighting the nuanced
impacts of economic crises on regional disparities. The integration of
transport infrastructure and education into this discourse adds an
additional layer of complexity, underscoring its potential as a catalyst
for regional development and equality.

\hypertarget{data-and-descriptive}{%
\section{Data and Descriptive}\label{data-and-descriptive}}

Trough Eurostat, we download the datasets nama\_10r\_3gdp ,
demo\_r\_pjanggr3, demo\_r\_mlifexp, tran\_r\_net and edat\_lfse\_04 as
csv files. Selecting our countries of study and timeperiod. Selecting
the required NUTS level and specified the data to be in million Euro.

\hypertarget{gdp}{%
\subsection{GDP}\label{gdp}}

The GDP dataset provides insights into GDP at regional level using the
NUTS classification system. It furnishes GDP values in both current
prices and adjusted for inflation, with figures given in purchasing
power standards (PPS) to account for price level differences between
countries. The data is mostly structured by year and region
\emph{Database - Eurostat} (n.d.).

The GDP at market prices represents the final result of production
activities of resident producer units within a region or nation. It is
calculated as the sum of the gross value added across various
institutional sectors or industries. Furthermore, augmented by taxes and
reduced by subsidies on products (which are not allocated to specific
sectors or industries) \emph{Database - Eurostat} (n.d.). This also
balances out in the total economy production account. In terms of
methodology, while national accounts compile GDP from the expenditure
side, regional accounts don't adopt this approach due to the
complexities of accurately mapping inter-regional flows of goods and
services.

The different measures for the regional GDP are absolute figures in €
and Purchasing Power Standards (PPS), figures per inhabitant and
relative data compared to the EU Member States average \emph{Database -
Eurostat} (n.d.).

\hypertarget{population}{%
\subsection{Population}\label{population}}

Eurostat's records annual population data with NUTS classification. Our
dataset includes information on births, deaths, net migration, and may
also include demographic information on age and gender. Displayed in a
year-by-region format, with yearly interval updates \emph{Database -
Eurostat} (n.d.).

Eurostat's primary source for yearly demographic data at the regional
level stems from the Unified Demography (Unidemo) project. The project
covers 37 countries and is the central repository for demographic and
migration-related data \emph{Database - Eurostat} (n.d.). As well as,
specific metrics gathered under UNIDEMO encompass population counts at
the close of the calendar year and events such as births and deaths
occurring within that year \emph{Database - Eurostat} (n.d.).
Additionally, data on marriages, divorces, and migration flows are
recorded .

For the purpose of this research, the demographic data references the
NUTS 2016 classification, which provides a detailed breakdown of the
European Union's territory \emph{Database - Eurostat} (n.d.).

\hypertarget{demographic-nuts-2}{%
\subsection{Demographic (Nuts 2)}\label{demographic-nuts-2}}

Statistics on population change (demo\_r\_mlifexp) and~on population
structure was collected from Euro Stat (\emph{Database - Eurostat},
n.d.). Used to monitor demographic behavior within political, economic,
social and cultural contexts at NUTS 2 level. Same policy and
demographic classifications as previous population data set applies.

\hypertarget{transportation-infrastructure-nuts-2}{%
\subsection{Transportation Infrastructure (Nuts
2)}\label{transportation-infrastructure-nuts-2}}

This data-set (tran\_r\_net) presents recent data on the inland
transport network of the European Union (EU), EFTA and candidate
countries presenting motorways, railways and inland waterways
(\emph{Database - Eurostat}, n.d.). Furthermore used as a varable of of
interest (MLR) in our section growth and inequality. Moreover, to
excamine if evolution of the transport network is closely linked to the
general development of the economy.

The collected data are compiled by the competent national authorities.
Reported annually be the National Statistical Office, the Civil Aviation
Authorities (air transport), the Ministry of Transport (inland
waterways, railway and road networks, victims), and the National
Maritime Administration (maritime transport) (\emph{Database -
Eurostat}, n.d.).

\hypertarget{education-nuts-2}{%
\subsection{Education (Nuts 2)}\label{education-nuts-2}}

This data set (edat\_lfse\_04) contains population by educational
attainment level, sex and NUTS 2 regions (\%). Including data on the
highest level of education successfully completed by the individuals of
a given population. Furthermore, data on young people neither in
employment nor in education and training -- NEET, early leavers from
education and training and the labor status of young people by years
since completion of highest level of education. Calculated as annual
averages of quarterly EU Labor Force Survey data (EU-LFS).
